Canary in the Coal Mine: Anti-NGO Legislation as a Predictor of Civil Society Repression

### 2021-04-04 manuscript/manuscript.Rmd
PHEWWWW ALL READY FOR ISA

---
title: "Canary in the Coal Mine: Anti-NGO Legislation as a Predictor of Civil Society Repression"
short-title: "Canary in the Coal Mine"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "March 23, 2021"
published: This is an early draft of a working paper.
git-repo: https://github.com/andrewheiss/canary-ngos
abstract: >-
  Many countries across the globe are increasingly cracking down on civil society groups through legal restrictions. While research on democratic backsliding has tended to focus on rights and institutions in the electoral arena, this new wave of laws—many of which appear mundane on their face—represent a bureaucratic form of repressive that could be indicative of more severe human rights abuses and backsliding. This is especially the case for electoral democracies, which unlike autocracies do not aggressively and openly attack civic space. In this paper, we ask: do legal crackdowns on NGOs predict broader civil society repression? We argue that this is the case because anti-NGO laws are among the most subtle means of repression and attract the least least domestic and international condemnation, as many audiences mistake this repression as regulation and are not personally affected by it. These laws also make it easier for states to prevent and deter future large-scale popular mobilization. Using original data on civil society restrictions over the last three decades, we test whether NGO crackdown is a predictor of attacks on journalists, independent media, academic freedom, and freedoms of expression and association. This project has important implications: if crackdown on NGOs is indeed an early warning, it can potentially prevent further deteriorations in space for civil society. 
keywords: "civil society restrictions, human rights, journalists"
reference-section-title: References
bibliography: bibliography.bib
csl: pandoc/csl/chicago-author-date.csl
title-page: true
editor_options: 
  chunk_output_type: console
---

@@

# R options
options(
  width = 90,  # Output width
  dplyr.summarise.inform = FALSE,  # Turn off dplyr's summarize() auto messages
  knitr.kable.NA = "",  # Make NAs blank in kables
  kableExtra.latex.load_packages = FALSE,  # Don't add LaTeX preamble stuff
  modelsummary_default = table_format  # Set modelsummary backend
)

library(tidyverse)
library(here)

source(here("lib", "graphics.R"))
```

Governments across the globe have been cracking down on civil society organizations (CSOs) -- a phenomenon referred to as closing civic space or shrinking civic space (CITE). The non-violent nature of this repression is particularly notable. The last two decades have seen a proliferation of laws designed to limit the influence of civil society organizations, with more than 100 countries obstructing, repressing, or closing the legal environment for civil society. Since 2013, 91 countries had proposed or enacted 244 measures restricting civil society [@icnl:2021]. In Human Rights Watch's 2016 World Report, Executive Director Kenneth Roth argued that civil society was under more aggressive attack than at any time in recent memory, and over the past decade, the majority of the world has seen a substantial narrowing of allowable civic space. As seen in Figure \@ref(fig:civicus-map), by the end of 2020, only `r civicus$Open$prop_nice` of countries had open and unrestricted civil societies [@civicusmonitor; @ChaudhryHeiss:2021].

WHAT KINDS OF LAWS, WHAT DO THESE LAWS DO.

The use of legal tools instead of violent repression means that NGOs have to adapt their practices to stay within the bounds of permitted activities and operations. While in some cases this has led to the depoliticization of NGOs and avoidance of direct political activism by CSOs, states cannot soften or neutralize all forms of political activities in subtle, legal ways that they have with NGOs. Soon or later, states need to deal with threats in a more violent, confrontational manner. Studying these NGO laws can thus be the canary in the coal mine which signals the states that intend to continue down this repressive path. What does state targeting of NGOs tell us about a state's broader respect for civil society and the overall level of human rights repression? Can the crackdown on NGOs act as an indicator for subsequent increasing repression?

ARGUMENT, METHODS, AND FINDINGS. 
In this paper, we argue that information about NGO crackdown can be a valuable indicator of human rights trajectories. Such crackdown can act as an early warning signal for worsening repression, suggesting which countries are testing the waters with comparatively 'milder' forms of repression before enacting more violent forms of crackdown. To assess if attacks on NGOs and activists can provide us useful information about future repression, we use an original dataset on state crackdown on NGOs using legal repression across all countries from 1990-2013. WHAT ELSE DO WE USE. Our analysis shows FILL

Our paper makes an important contribution in predicting (IS THIS THE RIGHT WORD) the onset of repressive state policies. While structural factors such as levels of democracy, economic development, electoral competitiveness can identify countries that are likely to have more or less respect for civic freedoms, these factors are also slow to change over time. As such, they may not be particularly suitable for forecasting short-term trajectories, especially for moderatively repressive countries. On the other hand, state repression of NGOs is more dynamic and provides information about changing levels of human rights standards and repression in subsequent years.

ROADMAP

# Assessing Changes in Government Repression

Repression is the use of coercive action against an individual or organization, within the territorial jurisdiction of the state, for the purpose of imposing a cost on the target as well as deterring specific activities and/or beliefs perceived to be challenging to government personnel, practices, or institutions (Goldstein 1978, xxvii; Poe and Tate 1994; Davenport 2007b; and Ritter and Conrad 2016).

One part of the existing literature focuses on the role of structural factors such as levels of democracy, economic development, electoral competitiveness in predisposing a state to increased use of repression. Most notably, both democracies and democratization is found to have a pacifying effect on state repression (Davenport and Armstrong 2004, Krain 1997, Richards 1999). Democracies are found to have a negative effect on repression largely because of the increased probability of sanctions against authorities for undesirable behavior (Davenport 2007, 41).

Another strand of research focuses on how the behavior of various actors shapes government response. One of the most consistent findings in the political violence literature is that authorities repress to counter ongoing challenges to the status quo. In other words, dissent increases repressive behavior (Davenport 2007).  However, dissent and repression have an endogenous relationship - state authorities also frequently engage in preventive repression to undermine or restrict groups’ abilities to dissent (Ritter and Conrad 2016; and Conrad and Ritter 2019.) In addition to attacking groups’ willingness to organize against states, such
preventive repression also damages their capacity to impose costs (Davenport 2007a, 47.)

NGOs are one actor that can pose challenges to political authorities, and state repression of these actors, especially through legal restrictions or administrative crackdown can also be considered a form of preventive repression. Below, we explore why government repress NGOs, especially through subtle, legal means.

## State repression of NGOs

A robust, strong civil society is widely considered as an important component on democratic transition and consolidation (Putnam). Politically, as important civil society actors, NGOs enable citizens to voice their grievances, seek redress, and carry out lobbying and advocacy activities to further their interests. They help keep government accountable, monitor human rights violations and corruption. From an economic perspective, NGOs provide services when the state falls short.

As a result of these roles, NGOs can wield real influence, which can be threatening to states. State crackdown on NGOs has become pervasive as these organizations have the ability to influence electoral politics (Dupuy, Ron, and Prakash 2016, Weiss 2009), aid mobilization (Murdie and Bhasin 2011; Bell et al. 2013; and Boulding 2014), and threaten a state’s economic interests (Dietrich and Murdie 2017; Lebovic and Voeten 2009; and Nielsen 2013.).

Why might repressive states start with non-violent forms of repression?
WHY THE GROWTH IN THESE LAWS
As someone points out, states have toned down 'open, violent confrontation' and favor the development of a 'civilized' oppression.

Leaders want to avoid accountability. Shape the narrative of their performance both domestically and internationally. Don't like being shamed.
Can be the beginning of a deteriorating human rights situation.

In the aftermath of the failed coup attempt in Turkey in 2016, the government initiated a widespread crackdown on civil society, leading to the closure of almost 1500 NGOs soon after the group. However, this crackdown, ``spiraled far beyond those with any possible connection to the coup itself." In October 2017, the government levied terrorism charges against human rights activists, including on Amnesty International's Turkey Director and Chair.  The government also imprisoned an increasing number of journalists in 2017 and 2018, and issued arrest warrants for numerous academics in 2018.

##Expectations

#Methods

#Results and Analysis

#Conclusion and Implications

>>>>

---
title: "Canary in the Coal Mine: Anti-NGO Legislation as a Predictor of Civil Society Repression"
short-title: "Canary in the Coal Mine"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "April 3, 2021"
published: This is an early draft of a working paper.
git-repo: https://github.com/andrewheiss/canary-ngos
abstract: >-
  Many countries across the globe are increasingly cracking down on civil society groups through legal restrictions. This new wave of attempts to repress civil society organization-many of which are sanctioned by law—represent a bureaucratic form of repression that could indicate more severe human rights abuses and democratic backsliding. This is especially the case for electoral democracies, which unlike autocracies do not aggressively and openly attack civic space. In this paper, we ask: do crackdowns on NGOs predict broader human rights repression? We argue that this is the case because anti-NGO laws are among the most subtle means of repression and attract the least least domestic and international condemnation, as many audiences mistake this repression as regulation and are not personally affected by it. These laws also make it easier for states to prevent and deter future large-scale popular mobilization. Using original data on civil society restrictions over the last three decades, we test whether NGO crackdown is a predictor of political terror, physical violence, and violations of private civil liberties. This project has important implications: if crackdown on NGOs is indeed an early warning, it can potentially prevent further deteriorations in space for civil society. 
keywords: "civil society restrictions, human rights, political terror, civil liberties"
reference-section-title: References
bibliography: bibliography.bib
title-page: true
editor_options: 
  chunk_output_type: console
---

@@

# R options
options(
  width = 90,  # Output width
  dplyr.summarise.inform = FALSE,  # Turn off dplyr's summarize() auto messages
  knitr.kable.NA = "",  # Make NAs blank in kables
  kableExtra.latex.load_packages = FALSE,  # Don't add LaTeX preamble stuff
  modelsummary_default = table_format  # Set modelsummary backend
)
```

```{r libraries-data, include=FALSE}
library(tidyverse)
library(targets)
library(kableExtra)
library(modelsummary)
library(patchwork)
library(scales)
library(ggtext)
library(sf)
library(here)
library(brms)

withr::with_dir(here::here(), {
  source(tar_read(plot_funs))

  canary <- tar_read(panel)
  tar_load(civicus_map_data)
  tar_load(civicus_clean)

  tar_load(c(m_pts_total, m_pts_advocacy, m_pts_entry, m_pts_funding,
             m_clphy_total, m_clphy_advocacy, m_clphy_entry, m_clphy_funding,
             m_clpriv_total, m_clpriv_advocacy, m_clpriv_entry, m_clpriv_funding,
             m_pts_v2csreprss, m_clphy_v2csreprss, m_clpriv_v2csreprss))

  # Model tables
  tar_load(c(models_paper_pts, models_paper_clphy, models_paper_clpriv))

  # Load lookup list for coefficients in regression tables
  tar_load(coef_list)
})

# Temporary fake cache stuff
mfx_e1a <- readRDS(here("temp_cache", "mfx_e1a.rds"))
mfx_e1b <- readRDS(here("temp_cache", "mfx_e1b.rds"))
mfx_e1c <- readRDS(here("temp_cache", "mfx_e1c.rds"))
df_mfx_e2a <- readRDS(here("temp_cache", "df_mfx_e2a.rds"))
df_mfx_e2b <- readRDS(here("temp_cache", "df_mfx_e2b.rds"))
df_mfx_e2c <- readRDS(here("temp_cache", "df_mfx_e2c.rds"))
```

```{r calc-civicus}
civicus <- civicus_clean %>%
  count(category) %>% 
  mutate(prop = n / sum(n)) %>% 
  mutate(prop_nice = percent_format(accuracy = 1)(prop)) %>% 
  split(.$category)
```

Governments across the globe have been cracking down on civil society organizations (CSOs), a phenomenon commonly referred to the closing or shrinking of civic space [@CarothersBrechenmacher:2014; @DPR:2016; @Chaudhry:2016; @DupuyPrakash:2018]. The last two decades have seen a proliferation of steps designed to limit the influence of CSOs and non-governmental organizations (NGOs), and more than 100 countries have proposed or enacted 244 measures restricting, repressing, or shutting down civil society since 2013 [@icnl:2021]. In Human Rights Watch's 2016 World Report, Executive Director Kenneth Roth argued that civil society was under more aggressive attack than at any time in recent memory, and over the past decade, the majority of the world has seen a substantial narrowing of allowable civic space. As seen in Figure \@ref(fig:civicus-map), by the end of 2020, only `r civicus$Open$prop_nice` of countries had open and unrestricted civil societies [@civicusmonitor; @ChaudhryHeiss:2021].

```{r civicus-map, fig.cap="2020 CIVICUS Monitor civic space ratings"}
ggplot() +
  geom_sf(data = civicus_map_data, aes(fill = fct_rev(category)), size = 0.15, color = "black") +
  coord_sf(crs = st_crs("ESRI:54030"), datum = NA) +  # Robinson
  scale_fill_viridis_d(option = "plasma", end = 0.9, 
                       na.translate = FALSE, name = "Civic space") +
  theme_void(base_size = 9, base_family = "Inter Light") +
  theme(legend.position = "bottom",
        legend.key.size = unit(0.7, "lines"))
```

This hostility to civil society is not unique to autocratic governments or even illiberal democracies [@AbramowitzSchenkkan:2018]—long-established liberal democracies have also increased restrictions on civil society in their countries. Over the past decade, India has canceled the licenses of tens of thousands of NGOs, targeted major groups such as Greenpeace International and Ford Foundation, and has banned 5,000 NGOs from receiving foreign funds [@press-trust-india:2015; @press-trust-india:2019]. In 2013 Canada attempted to revoke the charitable status of civil society organizations for opposing the Northern Gateway pipeline and Canadian mining companies' initiatives abroad [@CIVICUS:2013]. In 2019, activists from the US-based nonprofit No More Deaths were arrested for providing food and water to migrants a the US-Mexico border and faced a 20-year prison sentence before their convictions were overturned in 2020 [@Cramer:2020].

Anti-civil society measures are occasionally violent, but notably, government repression of NGOs is generally non-violent. Rather than publicly arrest or beat activists, states implement "administrative crackdown," or the passage of laws to create barriers to NGO advocacy, entry, funding [@Chaudhry:2016]. The use of legal tools instead of violent repression means that NGOs have to adapt their practices to stay within the bounds of permitted activities and operations [@Heiss:2019; @Heiss:2017]. While in some cases this has led to the depoliticization of NGOs and avoidance of direct political activism by CSOs, states cannot soften or neutralize all forms of political activities in subtle, legal ways that they have with NGOs. Sooner or later, states need to deal with threats to their stability in a more direct, violent, confrontational manner. Administrative NGO laws and general civil society repression could be an important early warning sign—or a canary in the coal mine—of worsening human rights repression in a state, signaling that a state will continue down the path of further repression. What does state targeting of NGOs tell us about a state's broader respect for civil society and the overall level of human rights repression?  Can the crackdown on NGOs act as an indicator for subsequent increasing repression?

In this paper, we explore the power of anti-NGO repression in predicting human rights abuses, political terror, physical violence, and violations of private civil liberties. We argue that information about NGO crackdown can be a valuable indicator of human rights trajectories. In some cases, such crackdown can act as an early warning signal for worsening repression, suggesting which countries are testing the waters with comparatively "milder" methods of crackdown before enacting more violent forms of repression. To assess if attacks on NGOs and activists can provide us useful information about future repression, we use an original dataset on state crackdown on NGOs using legal repression across all countries from 1990–2013. We differentiate between two forms of civil society repression: official de jure anti-NGO restrictions that are codified in law, and general de facto levels of civil society repression. We propose and test two empirical expectations—that de jure and de factor restrictions predict future political terror, physical violence, and government encroachment on private civil liberties. Using multilevel Bayesian modeling, we find that formal de jure laws provide very little additional information when attempting to predict future repression. However, we find that accounting for more general civil society repression does boost our models' predictive power. Increasing de facto civil society repression predicts an increase in the probability of moderate political terror and results in a 3% decrease in indexes of physical violence and private civil liberties.

Our paper makes an important contribution in predicting the onset of repressive state policies. While structural factors such as levels of democracy, economic development, electoral competitiveness can identify countries that are likely to have more or less respect for civic freedoms, these factors are also slow to change over time. As such, they may not be particularly suitable for forecasting short-term trajectories, especially for moderately repressive countries. On the other hand, state repression of NGOs is more dynamic and provides information about changing levels of human rights standards and repression in subsequent years.

Below, we first look at existing literature on repression and political violence, outlining state motivations to repress NGOs. We then elaborate on why attacking NGOs may be an important predictor of subsequent repression. Using both original data as well as data from several existing datasets, we build an ensemble of predictive models and explore (1) the marginal effects of civil society repression on different forms of broader human rights repression, and (2) the out-of-sample predictive performance of each model. We conclude with a discussion of our findings and their implications for assessing the repression of human rights.

# Government repression and NGO restrictions

Repression is the use of coercive action against an individual or organization, within the territorial jurisdiction of the state, for the purpose of imposing a cost on the target as well as deterring specific activities and/or beliefs perceived to be challenging to government personnel, practices, or institutions [@Goldstein:1978, p. xxvii; @PoeTate:1994; @Davenport:2007; and @RitterConrad:2016].

One part of the existing literature focuses on the role of structural factors such as levels of democracy, economic development, electoral competitiveness in predisposing a state to increased use of repression. Most notably, democratic forms of government—and democratization in general—has a pacifying effect on state repression [@DavenportArmstrong:2004; @Krain:1997; @Richards:1999]. Democracies have a negative effect on repression largely because of the increased probability of sanctions against authorities for undesirable behavior [@Davenport:2007a, 47].

Another strand of research focuses on how the behavior of various actors shapes government response. One of the most consistent findings in the political violence literature is that authorities repress to counter ongoing challenges to the status quo. In other words, dissent increases repressive behavior [@Davenport:2007]. However, dissent and repression have an endogenous relationship: state authorities also frequently engage in preventive repression to undermine or restrict groups' abilities to dissent [@RitterConrad:2016; @ConradRitter:2019]. In addition to attacking groups willingness to organize against states, such preventive repression also damages states' capacity to impose costs [@Davenport:2007a, 47].

NGOs are one actor that can pose challenges to political authorities, and state repression of these actors, especially through legal restrictions or administrative crackdown can also be considered a form of preventive repression. A robust, strong civil society is widely considered as an important component on democratic transition and consolidation [@Putnam:1994]. Politically, as important civil society actors, NGOs enable citizens to voice their grievances, seek redress, and carry out lobbying and advocacy activities to further their interests. They help keep government accountable, monitor human rights violations and corruption. From an economic perspective, NGOs provide services when the state falls short. As a result of these roles, NGOs can wield real influence, which can be threatening to states. State crackdown on NGOs has become pervasive as these organizations have the ability to influence electoral politics [@DPR:2016; @Weiss:2009], aid mobilization [@MurdieBhasin:2011; @BBCM2014; @Boulding:2014], and threaten a state's economic interests [@DietrichMurdie:2017; @LebovicVoeten:2009; @Nielsen:2013].

There is no doubt that violence is often effective at curbing dissent. But it can also backfire, leading to widespread protests, decreasing leaders' legitimacy and increasing their criminal liability [@KimSikkink:2010; @Sikkink:2011]. In contrast to violence, administrative crackdown has fewer adverse consequences—citizens may see it as regulation rather than repression, resulting in little to no counter-mobilization. It may also circumvent concerns among state security or bureaucratic agents about illegitimate orders based on violent repression [@StephanChenoweth:2008]. There are also fewer international consequences since laws—unlike violence—rarely invite condemnation or threats of aid withdrawal [@CarothersBrechenmacher:2014].

We argue that crackdown on NGOs foreshadows future government transgressions. Autocratic leaders want to avoid accountability, and NGOs often stand in conflict with that goal. Civil society organizations, including NGOs, can help shape the narrative of leaders' performance both domestically and internationally. Monitoring by NGOs and subsequent shaming can include significant external revenue loss to states in the form of loss of aid, trade, and foreign direct investments [@KeckSikkink:1998; @BarryClayFlynn:2013]. Government attempts to repress NGOs can signal the beginning of a deteriorating human rights situation. Such repression reflects the government's willingness and capability to use measures that can make it avoid accountability. Once a state targets NGOs, over time it may grow less hesitant to apply repression more widely and more severely.

In exploring the impact of state crackdown on NGOs on subsequent human rights conditions, we look at both *de jure* measures—or formal legislation that seeks to repress NGOs—as well as the *de facto* implementation of these measures. This distinction is important. While states may pass anti-NGO laws, these laws can also lie dormant. For instance, Russia passed its 2015 Undesirable Organizations Law with the purpose of targeting and eliminating specific foreign-connected NGOs it deemed dangerous. One of the law's sponsors, Aleksandr Tarnavsky, described the law as a preventative measure that would not affect the majority of NGOs working in Russia. Rather, the law would be a "weapon hanging on the wall that never fires" and stand as a warning to potentially uncooperative NGOs [@Kozenko:2015].

Simply noting the passage of (or counting) these laws may therefore miss the effect of their de facto implementation. Moreover, the implementation of these laws may not be limited to its legal character. The government may also engage in harassment of NGOs and activists to dissuade NGOs from operation or expressing themselves. This may include measures such as detentions, short-term incarceration, as well as beatings, threats to families and destruction of valuable property. Finally, counting official laws can also undermeasure civil society repression because governments can repress NGOs without passing or relying on formal laws. As prior research in Egypt, Ethiopia, and Russia points out, "rather than systematically enforcing all repressive legal measures-often impossible owing to capacity constraints," governments may choose to use extra-legal means to intimidate, harass or prosecute a few NGOs or activists. These cases, especially during moments or crises or contention, "highlight to the wider NGO community-and citizenry-that the government is willing and capable of using the law to repress dissent" [@Brechenmacher:2017, 94].

Based on these purposes of civil society restrictions and the different methods states use to implement them, we propose two empirical expectations about the relationship between civil society repression and general human rights:

- **Empirical expectation 1**: If a country imposes a formal anti-NGO law, then it is likely to signal future deteriorating human rights conditions.
- **Empirical expectation 2**: If a country represses civil society, then it is likely to signal future deteriorating human rights conditions.

# Data and modeling approach

To test these empirical expectations, we use a new dataset of civil society laws, as well as measures from several other larger datasets focused on human rights, political terror, and democratization. Table \@ref(tab:var-summary) provides a summary of our key variables.

```{r var-summary}
tribble(
  ~Measure, ~Category, ~Description, ~Source,
  "Political Terror Scale (PTS)", NA, "Ordered category ranging from 1 (least terror) to 5 (most terror)", "Gibney et al. (2019)",
  "Physical violence index", NA, "0–1; higher values indicate less violence", "Coppedge et al. (2020); v2x_clphy",
  "Private civil liberties index", NA, "0–1; higher values indicate greater respect for civil liberties", "Coppedge et al. (2020); v2x_clpriv",
  "Total legal barriers", "De jure laws", "Count of all anti-NGO legal barriers", "Chaudhry (2016)",
  "Barriers to advocacy", "De jure laws", "Count of laws restricting NGO advocacy", "Chaudhry (2016)",
  "Barriers to entry", "De jure laws", "Count of laws restricting NGO entry, including registration", "Chaudhry (2016)",
  "Barriers to funding", "De jure laws", "Count of laws restricting NGO funding sources", "Chaudhry (2016)",
  "CSO repression", "De facto repression", "≈ −3–3; higher values indicate less repression", "Coppedge et al. (2020); v2csreprss"
) %>%
  kbl(caption = "Outcomes and predictors",
      booktabs = TRUE, linesep = "\\addlinespace") %>% 
  kable_styling(font_size = 8) %>% 
  row_spec(0, bold = TRUE) %>% 
  column_spec(1, width = "0.14\\\\textwidth") %>% 
  column_spec(2, width = "0.14\\\\textwidth") %>% 
  column_spec(3, width = "0.4\\\\textwidth") %>% 
  column_spec(4, width = "0.2\\\\textwidth") %>% 
  pack_rows(index = c("Outcomes" = 3, "Predictors" = 5))
```

```{r var-summary1, eval=FALSE, include=FALSE}
tribble(
  ~Measure, ~Category, ~Description, ~Source,
  "Political Terror Scale (PTS)", NA, "Ordered category ranging from 1 (least terror) to 5 (most terror)", "@GibneyCornettWood:2019",
  "Physical violence index", NA, "0–1; higher values indicate less violence", "@vdem-v10; v2x_clphy",
  "Private civil liberties index", NA, "0–1; higher values indicate greater respect for civil liberties", "@vdem-v10; v2x_clpriv",
  "Total legal barriers", "De jure laws", "Count of all anti-NGO legal barriers", "\\@Chaudhry:2016",
  "Barriers to advocacy", "De jure laws", "Count of laws restricting NGO advocacy", "\\@Chaudhry:2016",
  "Barriers to entry", "De jure laws", "Count of laws restricting NGO entry, including registration", "@Chaudhry:2016",
  "Barriers to funding", "De jure laws", "Count of laws restricting NGO funding sources", "@Chaudhry:2016",
  "CSO repression", "De facto repression", "≈ −3–3; higher values indicate less repression", "@vdem-v10; v2csreprss"
) %>%
  kbl(caption = "Outcomes and predictors",
      booktabs = TRUE, linesep = "") %>% 
  kable_styling(font_size = 8.5) %>% 
  row_spec(0, bold = TRUE) %>% 
  column_spec(1, width = "0.14\\\\textwidth") %>% 
  column_spec(2, width = "0.14\\\\textwidth") %>% 
  column_spec(3, width = "0.5\\\\textwidth") %>% 
  column_spec(4, width = "0.2\\\\textwidth") %>% 
  pack_rows(index = c("Outcomes" = 3, "Predictors" = 5))
```

## Measuring de jure and de facto civil society repression

@christensen2013 collected data on formal NGO regulations in 90 countries to explore the determinants of laws that restrict NGO foreign funding laws specifically. @Chaudhry:2016 expanded this data so that it now includes official NGO laws for 162 countries between 1990 and 2013. We create indexes of the total number of laws passed in each country based on three broad categories, and we also sum each of these indexes to create a count of total legal barriers:

- Barriers to advocacy (2 possible barriers): NGOs are restricted from engaging in political activities, and advocacy restrictions differ if the NGO receives foreign funding
- Barriers to entry (3 possible barriers): registration is burdensome, NGOs are not allowed to appeal if denied registration, and entry requirements differ if the NGO receives foreign funding
- Barriers to funding (5 possible barriers): NGOs need prior approval to receive foreign funding, NGOS are required to channel foreign funding to state-owned banks, NGOS face additional restrictions if receiving foreign funding, and NGOs are prohibited from receiving foreign funding

```{r barriers, fig.width=7, fig.height=6, fig.asp=NULL, fig.cap="Number of legal barriers to NGO activity per country over time (de jure legislation) and average level of CSO repression across democracies and non-democracies (de facto implementation)"}
regulations <- tribble(
    ~question, ~barrier,       ~question_clean,                  ~ignore_in_index, ~question_display,
    "q1a",     "association",  "const_assoc",                    TRUE,             "Constitutional associational rights",
    "q1b",     "association",  "political_parties",              TRUE,             "Citizens form political parties",
    "q4b",     "advocacy",     "ngo_politics_intimidation",      TRUE,             "NGOs intimidated from politics",
    "q4c",     "advocacy",     "ngo_politics_foreign_fund",      FALSE,            "Political barriers differentXXXif foreign funds involved",
    "q2a",     "entry",        "ngo_register",                   TRUE,             "NGO registration required",
    "q2b",     "entry",        "ngo_register_burden",            FALSE,            "NGO registration burdensome",
    "q2c",     "entry",        "ngo_register_appeal",            FALSE,            "NGO registration appealXXXnot allowed",
    "q2d",     "entry",        "ngo_barrier_foreign_funds",      FALSE,            "Registration barriers differentXXXif foreign funds involved",
    "q3a",     "funding",      "ngo_disclose_funds",             TRUE,             "Funds must be disclosed",
    "q3b",     "funding",      "ngo_foreign_fund_approval",      FALSE,            "Prior approval requiredXXXfor foreign funds",
    "q3c",     "funding",      "ngo_foreign_fund_channel",       FALSE,            "Foreign funds channeledXXXthrough government",
    "q3d",     "funding",      "ngo_foreign_fund_restrict",      FALSE,            "Foreign funds restricted",
    "q3e",     "funding",      "ngo_foreign_fund_prohibit",      FALSE,            "Foreign funds prohibited",
    "q3f",     "funding",      "ngo_type_foreign_fund_prohibit", FALSE,            "Foreign funds prohibitedXXXfor some types of NGOs",
    "q4a",     "advocacy",     "ngo_politics",                   FALSE,            "NGOs restricted from politics"
) %>% 
  filter(!ignore_in_index) %>% 
  mutate(barrier_group = paste0("Barriers to ", str_to_lower(barrier))) %>% 
  select(barrier_group, barrier = question_clean, barrier_display = question_display)

df_barriers <- canary %>% 
  group_by(gwcode, year) %>% 
  summarize(across(one_of(regulations$barrier), ~. > 0)) %>% 
  group_by(year) %>% 
  summarize(across(-gwcode, ~sum(., na.rm = TRUE))) %>% 
  ungroup() %>% 
  pivot_longer(names_to = "barrier", values_to = "value", -year) %>% 
  left_join(regulations, by = "barrier") %>% 
  mutate(barrier_display = str_replace(barrier_display, "XXX", "\n")) %>%
  arrange(desc(value)) %>% 
  mutate(barrier_display = fct_inorder(barrier_display, ordered = TRUE))

num_countries <- length(unique(canary$gwcode))

entry_plot <- ggplot(filter(df_barriers, 
                            barrier_group == "Barriers to entry"), 
                     aes(x = year, y = value, 
                         color = barrier_display,
                         linetype = barrier_display)) +
  geom_line(size = 0.75) +
  expand_limits(y = c(0, 90)) +
  scale_y_continuous(sec.axis = sec_axis(~ . / num_countries,
                                         labels = percent_format(accuracy = 1)),
                     expand = c(0, 0)) +
  scale_color_viridis_d(option = "plasma", end = 0.8, name = NULL) +
  scale_linetype_manual(values = c("solid", "solid", "21"), name = NULL) +
  coord_cartesian(xlim = c(1990, 2015), ylim = c(0, 90)) +
  guides(color = guide_legend(nrow = 2)) +
  labs(x = NULL, y = "Number of countries") +
  theme_ngo(9) + 
  theme(legend.justification = "left") +
  facet_wrap(vars(barrier_group))

funding_plot <- ggplot(filter(df_barriers, 
                              barrier_group == "Barriers to funding"), 
                       aes(x = year, y = value, 
                           color = barrier_display,
                           linetype = barrier_display)) +
  geom_line(size = 0.75) +
  scale_y_continuous(sec.axis = sec_axis(~ . / num_countries,
                                         labels = percent_format(accuracy = 1)),
                     expand = c(0, 0)) +
  scale_color_viridis_d(option = "plasma", end = 0.8, name = NULL) +
  scale_linetype_manual(values = c("solid", "solid", "solid", "21", "21"), name = NULL) +
  coord_cartesian(xlim = c(1990, 2015), ylim = c(0, 45)) +
  guides(color = guide_legend(nrow = 3),
         linetype = guide_legend(nrow = 3)) +
  labs(x = NULL, y = "Number of countries") +
  theme_ngo(9) + 
  theme(legend.justification = "left") +
  facet_wrap(vars(barrier_group))

advocacy_plot <- ggplot(filter(df_barriers, 
                               barrier_group == "Barriers to advocacy"), 
                        aes(x = year, y = value, 
                            color = barrier_display)) +
  geom_line(size = 0.75) +
  scale_y_continuous(sec.axis = sec_axis(~ . / num_countries,
                                         labels = percent_format(accuracy = 1)),
                     expand = c(0, 0)) +
  scale_color_viridis_d(option = "plasma", end = 0.8, name = NULL) +
  coord_cartesian(xlim = c(1990, 2015), ylim = c(0, 60)) +
  guides(color = guide_legend(nrow = 2)) +
  labs(x = NULL, y = "Number of countries") +
  theme_ngo(9) + 
  theme(legend.justification = "left") +
  facet_wrap(vars(barrier_group))

df_csreprss_plot <- canary %>%
  group_by(year, autocracy = v2x_regime <= 1) %>%
  nest() %>% 
  mutate(cis = data %>% map(~ mean_cl_normal(.$v2csreprss))) %>% 
  unnest(cis) %>% 
  ungroup() %>% 
  filter(!is.na(autocracy)) %>% 
  mutate(fake_facet_title = "CSO repression",
         autocracy = factor(autocracy, 
                            labels = c("Democracies (Regimes of the World > 1)",
                                       "Non-democracies (Regimes of the World ≤ 1)"), 
                            ordered = TRUE))

csreprss_plot <- ggplot(df_csreprss_plot, aes(x = year, y = y)) +
  geom_ribbon(aes(ymin = ymin, ymax = ymax, fill = autocracy), alpha = 0.2,) +
  geom_line(aes(color = autocracy), size = 0.75) +
  scale_color_viridis_d(option = "plasma", end = 0.8, name = NULL) +
  scale_fill_viridis_d(option = "plasma", end = 0.8, name = NULL) +
  coord_cartesian(xlim = c(1990, 2015)) +
  guides(color = guide_legend(nrow = 2)) +
  labs(y = "Average CSO repression", x = NULL) +
  theme_ngo(9) +
  theme(legend.justification = "left") +
  facet_wrap(vars(fake_facet_title))

barriers_summary <- 
  ((advocacy_plot | funding_plot) / 
     (funding_plot | csreprss_plot)) &
  theme(legend.text = element_text(size = rel(0.7)),
        axis.title.y = element_text(margin = margin(r = 3)),
        legend.box.margin = margin(t = -0.5, unit = "lines"))

barriers_summary
```

As mentioned previously, simply counting the number of NGO laws does not necessarily capture the restrictiveness of the general civil society environment in a country. Governments might repress civil society without formal laws or pass restrictive laws that they can keep in reserve for later emergencies. To measure the general level of civil society repression in a country, we use V-Dem's CSO repression indicator (`v2csreprss` in @vdem-v10), which captures the extent to which governments harass, deter, liquidate, and arrest members of civil society organizations. Figure \@ref(fig:barriers) shows each of these types of legal barriers per country over time, as well as average levels of civil society repression by regime type. While anti-NGO laws have increased steadily since 1990—particularly laws related to foreign funding—there has not necessarily been a corresponding rise in civil society repression, which has remained fairly constant across different regime types. This divergence is likely indicative of the split between de jure laws and their de facto implementation—a rise in laws is not necessarily tied to direct CSO repression.

## Measuring repression

We measure general human rights repression in three different ways. First, following @GohdesCarey:2017, who use the killing of journalists to predict worsening human rights conditions, we use the Political Terror Scale (PTS), which assigns countries to five different categories of repression each year [@GibneyCornettWood:2019], with Level 5 representing widespread political violence, including arrests, political murders, disappearances, and torture. As seen in the left panel of Figure \@ref(fig:repression), the number of countries at Level 5 has decreased since the 1990s, while most countries are coded as Levels 2 and 3. To supplement the PTS, we also turn to two indicators from V-Dem: a physical violence index (`v2x_clphy`) that measures freedom from torture and political killings specifically, and a private civil liberties index (`v2x_clpriv`) that measures respect for property rights, labor rights, freedom of religion, freedom of movement, and other non-political human rights. V-Dem includes several other rights indicators, including one for political civil liberties that captures many of the same dimensions as the PTS, but civil society repression is included as a factor for these other indexes, making it inappropriate to use as an outcome variable. The right panel of Figure \@ref(fig:repression) shows that the average level of these two rights indexes has improved over time.
